Концентрация юрфирм и доступность юруслуг: как они связаны?

### 2024-05-14 legal-companies-concentration-and-availability/journal-article/Article.Rmd
Modify plots, edit results and discussion, #2

```{r imports, include=FALSE, cache = TRUE}
library(here)
source(here("common", "Script.R"))
```

@@

**Data.** This research utilizes a geo-referenced dataset on small and medium-sized law firms incorporated in Russia. The dataset is created from FTS open data dumps (archives) of the state SMB Registry[^1] and FTS open data on the average number of employees of ogranizations[^2] using the Python command-line tool *ru-smb-companies*[^3] developed by the author of the paper. It contains information about the particular companies, including company's name, taxpayer ID, region and settlement of incorporation, main activity code. Law firms are selected by the main activity code equal to 69.10, and all the other companies were dropped from the source data. This technique of selection by the main activity code was derived from the analytical paper on the Russian legal services market prepared by the members of the Institute for the Rule of Law [@Moiseeva2016]. Additionally, the dataset was filtered by the following conditions: the year is 2021, the company is an organization rather than a sole entrepreneur, the company has declared non-zero revenue and expenditure for the target year (implying that the company is active), and the company's address of registration is in a city rather than in a rural area. During the exploratory data analysis, four cities (Innopolis, Kirovsk, Bronnytsy, Krasnoarmeisk) were identified as outliers based on the number of law firms or their employees per 100,000 population and were subsequently excluded from the data. In addition to the main dataset, a data on cities population freely distributed by the commercial company HFLabs is used [^4].

@@

@@

**Theoretical models.** Despite the perceived simplicity of the modelling described earlier, it has a theoretical mathematical background, and several possible types of outcome may be laid out right now before the analysis of the data in hand. It is known that a linear model is described by the two coefficients, intercept (a free term of the linear equation showing the relative shift of the regression line along the y axis) and slope (a coefficient before the independent variable showing the angle between the regression line and x axis). The slope can be either positive, negative, or zero. As a special edge case, it can be also infinite (i. e. regression line is vertical), even though, strictly speaking, such a dependency is not a function. Thus, the linear fit described earlier can yield a result of one of four types, each corresponding to one of the slope options and, subsequently, the pattern of relationship between the number of law firms and provision of legal services. These basic theoretical types are shown in Figure 1.

@@

Figure 1 shows the relationship between the concentration of legal companies and the availability of legal services, as measured by the metrics proposed in the Methods section. The plot displays the groups by city population and city rank within the region, in addition to the general trend. The overall connection between the concentration and availability appears to be linear. The linear model results demonstrate that the relationship between the two variables can be described by the formula *`r ca_model_formula`*, where both the intercept and slope are highly significant *(p < 0.001)*, and the explained variance is relatively high *(R^2^ = `r round(ca_model_summary$adj.r.squared, 2)`)*.

Examining the regression line common to all data points, we see that while the linearity is generally good and the linear modeling results are satisfactory, the total distribution is fan-shaped. This means that the variance of error of the linear model decreases as both the dependent and independent variables increase. However, this is mainly due to the presence of several distinct population-size groups of cities in the data. Within each of these groups, linear models satisfy the assumptions better. The link between concentration and availability is similar across all city population groups, and the regression lines are nearly parallel, indicating equal slope coefficients and differing only in intercepts. Additionally, it can be concluded that, in general, larger cities have higher concentrations and availability, although the relationship is relatively weak.

@@

In general, cities with the top rank by the population in their respective regions tend to have better availability of legal services. One possible reason for this is the considerably big size of such cities.

Moscow and Saint Petersburg are two remarkable outliers with extremely high concentration of legal services and a good level of legal services availability. Moscow alone accounts for almost one-third of all legal companies' employees, and Saint Petersburg accounts for about 10%. The position of the corresponding data points relative to the main regression lines suggests that there is an “availability saturation” level, where further concentration has little effect on availability.

A high concentration of legal companies, a large population, and the city's top rank by population within a region are positively associated with the availability of legal services. Conversely, larger cities require more legal companies to meet the needs of their residents. As a result, there are various ways to achieve high availability of legal services, ranging from being a business and service center with a high concentration of legal firms to being a small yet well-served city where a small number of legal firms meet the needs of the local community.

>>>>

```{r imports, include=FALSE, cache = TRUE}
library(here)
source(here("journal-article", "Script.R"))
```

@@

**Data.** This research utilizes a geo-referenced dataset on small and medium-sized law firms incorporated in Russia. The dataset is created from FTS open data dumps (archives) of the state SMB Registry[^1] and FTS open data on the average number of employees of ogranizations[^2] using the Python command-line tool *ru-smb-companies*[^3] developed by the author of the paper. It contains information about the particular companies, including company's name, taxpayer ID, region and settlement of incorporation, main activity code. Law firms are selected by the main activity code equal to 69.10, and all the other companies were dropped from the source data. This technique of selection by the main activity code was derived from the analytical paper on the Russian legal services market prepared by the members of the Institute for the Rule of Law [@Moiseeva2016]. Additionally, the dataset was filtered by the following conditions: the year is 2021, the company is an organization rather than a sole entrepreneur, the company has declared non-zero revenue and expenditure for the target year (implying that the company is active), and the company's address of registration is in a city rather than in a rural area. In addition to the main dataset, a data on cities population freely distributed by the commercial company HFLabs is used [^4].

@@

During the exploratory data analysis, six cities are identified as outliers subsequently excluded from the data. Two of these cities, Moscow and Saint Petersburg, are excluded due to the extremely high number of law firms employees: almost 18,866 and 5,550 respectively, while the next city has only less than 2,000 employees. The four other cities (Innopolis, Kirovsk, Bronnytsy, Krasnoarmeisk) are characterized by the very high proportion of law firms employees in city's population (from 38 per 10,000 in Krasnoarmeisk to 417 in Innopolis) compared to less than 20 per 10,000 in all the other cities.

@@

**Theoretical model.** Despite the perceived simplicity of the modelling described earlier, it has a theoretical mathematical background, and several possible types of outcome may be laid out right now before the analysis of the data in hand. It is known that a linear model is described by the two coefficients, intercept (a free term of the linear equation showing the relative shift of the regression line along the y axis) and slope (a coefficient before the independent variable showing the angle between the regression line and x axis). The slope can be either positive, negative, or zero. As a special edge case, it can be also infinite (i. e. regression line is vertical), even though, strictly speaking, such a dependency is not a function. Thus, the linear fit described earlier can yield a result of one of four types, each corresponding to one of the slope options and, subsequently, the pattern of relationship between the number of law firms and provision of legal services. These basic theoretical types are shown in Figure 1.

@@

Figure 1 shows the relationship between the concentration of legal companies and the availability of legal services, as measured by the metrics proposed in the Materials and methods section. The overall connection between the number of law firms and provision of legal services appears to be nonlinear and corresponds to a square root formula. However, the same relationship within each city size group is fitted well with a linear models. The coefficients of these models are different, and the slope decreases when the city population increases. While the regression lines for the small and medium cities are close to Type 4 (provision is independent of the number of lawyers because of the small variation in the number of lawers), the modelled relationship in millionaire cities resembles Type 1 (another option for independence, but due to relatively uniform and equal provision regardless the number of lawyers). The cities from big to extra-large are characterized by the positive relationship between the number of lawyers and the provision of legal services and can be classified as Type 2.

@@

In general, regional centers have relatively high number of lawyers and considerable level of legal services provision. This observation is probably explained by the centralization tendencies in the regions and the concentration of business and services in their administrative centers that situate the majority of government bodies and are also usually the most populous cities within a region.

Moscow and Saint Petersburg, though excluded from the data analysis and not shown on the main figure, are two remarkable outliers with extremely high concentration of legal services and a good level of legal services availability. Moscow alone accounts for almost one-third of all legal companies' employees, and Saint Petersburg accounts for about 10%. The position of the corresponding data points relative to the main regression lines suggests that there is an “availability saturation” level, where further concentration has little effect on availability.

The general tendencies observed in the Figure 1 should be supplemented by the analysis of the source data distribution shown in Figure 2 that extends the Figure 1 and displays the binned count of cities according to the number of lawyers and the provision of legal services. The overwhelming majority of cities is characterized by low level of both legal services provision and law firms development. Only a limited number (18) of cities is characterized by the high (more than 500) number of lawyers. It should be noted that high number of lawyers is good for legal services provision because all these cities are at least at the medium level of provision, and none of them has less than 5 lawyers per 10,000 people. However, cities with relatively low number of laywers can nevertheless reach top performance from the point of legal services provision that is on par with the cities with the highest number of lawyers. Moreover, it appears that such small yet well-served cities are even more typical than cities with high level of both lawyers and provision. The relative location of bin counts in the Figure 2 implies that there — besides the dominant bottom left corner — there are two main clusters of cities showing two different approaches of reaching high level of legal services provision. One approach is increase both in absolute number of law firms and relative number of lawyers per population probably charasteristic of the largest cities. Another is a relatively small yet sufficient number of lawyers typical probably for small or medium cities. Thus, apart from the general tendencies, the two-ways structure is found out.
